Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-04-04 paper.qmd
Updated the plot project

two_curvy_proj_true <- plot_proj(
  projected_df = projected_df, 
  model_df = true_model_df_proj, 
  axes = axes, 
  circle = circle, 
  point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
  plot_limits = c(-0.6, 0.6), 
  title = "a", 
  cex = 1.5, 
  axis_text_size = 4, 
  is_color = FALSE) +
  theme(aspect.ratio=1, 
        legend.position = "none")

@@

two_curvy_proj_tsne_first_model1 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "c", 
    cex = 1.5, 
    position = sc_ltr_pos,
    axis_text_size = 2, 
    is_color = FALSE) +
    theme(aspect.ratio = 1,
          legend.position = "none")

@@

projected_df <- proj_obj2$projected_df
model_df <- proj_obj2$model_df
axes <- proj_obj2$axes
circle <- proj_obj2$circle

@@

two_curvy_proj_true2 <- plot_proj(
    projected_df = projected_df, 
    model_df = true_model_df_proj, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
    line_param = c(0.5, 0.5, "black"), # linewidth, alpha
    plot_limits = c(-0.42, 0.42), 
    title = "b", 
    cex = 1.5, 
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

@@

two_curvy_proj_tsne_all_model1 <- plot_proj(
  projected_df = projected_df, 
  model_df = model_df, 
  axes = axes, 
  circle = circle, 
  point_param = c(0.1, 0.5, clr_choice), # size, alpha, color
  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
  plot_limits = c(-0.5, 0.5), 
  title = "d", 
  cex = 1, 
  position = sc_ltr_pos,
  axis_text_size = 2, 
  is_color = FALSE) #+
  #geom_segment(
  #  data = model_df, 
  #  aes(
  #    x = proj1_from, 
  #    y = proj2_from, 
  #    xend = proj1_to, 
  #    yend = proj2_to), 
  #  color = "#000000")

@@

two_curvy_proj_first_model1_umap <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "c2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(
        legend.position = "none"
    )

@@

two_curvy_proj_model_delaunay_umap <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df1, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "a2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

two_curvy_proj_model_benchmark2_umap <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df2, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "b2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

@@

projected_df <- proj_obj1$projected_df
model_df <- proj_obj1$model_df
axes2 <- proj_obj1$axes
circle <- proj_obj1$circle

## To add error category
projected_df <- projected_df |>
  mutate(cluster = factor(error_df_one_curvy_abs$error_cat_n2,
                          levels=c("deselected", "selected")))

five_c_shaped_proj_tsne_model1 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes2, 
    circle = circle, 
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    title = "a3", 
    cex = 1.5,
    position = c(0.08, 0.9),
    axis_text_size = 4, 
    is_color = TRUE) +
    scale_color_manual(values = c('#d9d9d9', '#800026')) +
    theme(aspect.ratio = 1,
          legend.position = "none")

projected_df <- projected_df |>
  select(-cluster) |>
  mutate(cluster = factor(error_df_one_curvy_abs$error_cat_n,
                          levels=c("deselected", "selected")))

five_c_shaped_proj_tsne_model1_selected <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes2, 
    circle = circle, 
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    title = "a3", 
    cex = 1.5,
    position = c(0.08, 0.9),
    axis_text_size = 4, 
    is_color = TRUE) +
    scale_color_manual(values = c('#d9d9d9', '#800026')) +
    theme(aspect.ratio = 1,
          legend.position = "none")
```

@@

pbmc_proj_umap_model1 <- plot_proj(
  projected_df = projected_df, 
  model_df = model_df, 
  axes = axes, 
  circle = circle, 
  point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
  plot_limits = c(-0.65, 0.65), 
  title = "a2", 
  cex = 2, 
  axis_text_size = 5,
  is_color = FALSE)

@@

projected_df <- proj_obj2$projected_df
model_df <- proj_obj2$model_df
axes <- proj_obj2$axes
circle <- proj_obj2$circle

pbmc_proj_umap_model2 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
    plot_limits = c(-1, 0.75), 
    title = "a3", 
    cex = 2, 
    axis_text_size = 5,
    is_color = FALSE)

@@

pbmc_proj_umap_model1_best <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
    plot_limits = c(-0.65, 0.65), 
    title = "b2", 
    cex = 2, 
    axis_text_size = 5,
    is_color = FALSE)

@@

projected_df <- proj_obj2$projected_df
model_df <- proj_obj2$model_df
axes <- proj_obj2$axes
circle <- proj_obj2$circle

pbmc_proj_umap_model2_best <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
    plot_limits = c(-1, 0.75), 
    title = "b3", 
    cex = 2, 
    axis_text_size = 5,
    is_color = FALSE)
```

@@

projected_df <- proj_obj1$projected_df |>
  mutate(cluster = tsne_minst_scaled$cluster)

model_df <- proj_obj1$model_df
axes <- proj_obj1$axes
circle <- proj_obj1$circle

mnist_proj_tsne_model1 <- plot_proj(
  projected_df = projected_df, 
  model_df = model_df, 
  axes = axes, 
  circle = circle, 
  point_param = c(0.5, 0.5), # size, alpha, color
  line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
  plot_limits = c(-1.1, 0.9), 
  title = "b", 
  cex = 1.5, 
  axis_text_size = 3, 
  is_color = TRUE) +
  scale_color_manual(values = c("#999999",'#ff7f00')) +
  theme(
    legend.position = "none"
  )

mnist_proj_tsne_model1_link <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(1.2, 0.5), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-1.1, 0.9), 
    title = "b", 
    cex = 2, 
    axis_text_size = 3, 
    is_color = TRUE) +
    scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
    theme(
        legend.position = "none"
    )

@@

projected_df <- proj_obj2$projected_df |>
  mutate(cluster = tsne_minst_scaled$cluster)

model_df <- proj_obj2$model_df
axes <- proj_obj2$axes
circle <- proj_obj2$circle

mnist_proj_tsne_model2 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.3), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-1, 1), 
    title = "c", 
    cex = 1.5, 
    axis_text_size = 3, 
    is_color = TRUE) +
    scale_color_manual(values = c("#999999",'#ff7f00')) +
    theme(
        legend.position = "none"
    )

mnist_proj_tsne_model2_link <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(1.2, 0.5), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-1, 1), 
    title = "c", 
    cex = 2, 
    axis_text_size = 3, 
    is_color = TRUE) +
    scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
    theme(
        legend.position = "none"
    )

>>>>

proj_obj1_true <- proj_obj1
proj_obj1_true[["model_df"]] <- true_model_df_proj

two_curvy_proj_true <- plot_proj(
  proj_obj = proj_obj1_true, 
  point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
  plot_limits = c(-0.6, 0.6), 
  title = "a", 
  cex = 1.5, 
  axis_text_size = 4, 
  is_color = FALSE) +
  theme(aspect.ratio=1, 
        legend.position = "none")

@@

proj_obj1[["axes"]] <- axes
proj_obj1[["circle"]] <- circle

two_curvy_proj_tsne_first_model1 <- plot_proj(
    proj_obj = proj_obj1, 
    point_param = c(0.5, 0.2, clr_choice), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "c", 
    cex = 1.5, 
    position = sc_ltr_pos,
    axis_text_size = 2, 
    is_color = FALSE) +
    theme(aspect.ratio = 1,
          legend.position = "none")

@@

@@

proj_obj2_true <- proj_obj2
proj_obj2_true[["model_df"]] <- true_model_df_proj

two_curvy_proj_true2 <- plot_proj(
    proj_obj = proj_obj2_true, 
    point_param = c(0.5, 0.5, clr_choice), # size, alpha, color
    line_param = c(0.5, 0.5, "black"), # linewidth, alpha
    plot_limits = c(-0.42, 0.42), 
    title = "b", 
    cex = 1.5, 
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

@@

proj_obj2[["axes"]] <- axes
proj_obj2[["circle"]] <- circle

two_curvy_proj_tsne_all_model1 <- plot_proj(
  proj_obj = proj_obj2, 
  point_param = c(0.1, 0.5, clr_choice), # size, alpha, color
  line_param = c(0.5, 0.5, "black"), # linewidth, alpha
  plot_limits = c(-0.5, 0.5), 
  title = "d", 
  cex = 1, 
  position = sc_ltr_pos,
  axis_text_size = 2, 
  is_color = FALSE) #+
  #geom_segment(
  #  data = model_df, 
  #  aes(
  #    x = proj1_from, 
  #    y = proj2_from, 
  #    xend = proj1_to, 
  #    yend = proj2_to), 
  #  color = "#000000")

@@

two_curvy_proj_first_model1_umap <- plot_proj(
    proj_obj = proj_obj1, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "c2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(
        legend.position = "none"
    )

@@

proj_obj1_temp1 <- proj_obj1
proj_obj1_temp2 <- proj_obj1
proj_obj1_temp1[["model_df"]] <- model_df1

two_curvy_proj_model_delaunay_umap <- plot_proj(
    proj_obj = proj_obj1_temp1, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "a2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

proj_obj1_temp2[["model_df"]] <- model_df2

two_curvy_proj_model_benchmark2_umap <- plot_proj(
    proj_obj = proj_obj1_temp2, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "b2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_color = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

@@

proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n2,
                          levels=c("deselected", "selected"))

five_c_shaped_proj_tsne_model1 <- plot_proj(
    proj_obj = proj_obj1,
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    title = "a3", 
    cex = 1.5,
    position = c(0.08, 0.9),
    axis_text_size = 4, 
    is_color = TRUE) +
    scale_color_manual(values = c('#d9d9d9', '#800026')) +
    theme(aspect.ratio = 1,
          legend.position = "none")

proj_obj1[["cluster"]] <- factor(error_df_one_curvy_abs$error_cat_n,
                          levels=c("deselected", "selected"))

five_c_shaped_proj_tsne_model1_selected <- plot_proj(
    proj_obj = proj_obj1, 
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    title = "a3", 
    cex = 1.5,
    position = c(0.08, 0.9),
    axis_text_size = 4, 
    is_color = TRUE) +
    scale_color_manual(values = c('#d9d9d9', '#800026')) +
    theme(aspect.ratio = 1,
          legend.position = "none")
```

@@

pbmc_proj_umap_model1 <- plot_proj(
  proj_obj = proj_obj1, 
  point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
  plot_limits = c(-0.65, 0.65), 
  title = "a2", 
  cex = 2, 
  axis_text_size = 5,
  is_color = FALSE)

@@

pbmc_proj_umap_model2 <- plot_proj(
    proj_obj = proj_obj2, 
    point_param = c(1.5, 0.2, "#a65628"), # size, alpha, color
    plot_limits = c(-1, 0.75), 
    title = "a3", 
    cex = 2, 
    axis_text_size = 5,
    is_color = FALSE)

@@

pbmc_proj_umap_model1_best <- plot_proj(
    proj_obj = proj_obj1,
    point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
    plot_limits = c(-0.65, 0.65), 
    title = "b2", 
    cex = 2, 
    axis_text_size = 5,
    is_color = FALSE)

@@

pbmc_proj_umap_model2_best <- plot_proj(
    proj_obj = proj_obj2,
    point_param = c(1.5, 0.2, "#999999"), # size, alpha, color
    plot_limits = c(-1, 0.75), 
    title = "b3", 
    cex = 2, 
    axis_text_size = 5,
    is_color = FALSE)
```

@@

proj_obj1[["cluster"]] <- tsne_minst_scaled$cluster

mnist_proj_tsne_model1 <- plot_proj(
  proj_obj = proj_obj1, 
  point_param = c(0.5, 0.5), # size, alpha, color
  line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
  plot_limits = c(-1.1, 0.9), 
  title = "b", 
  cex = 1.5, 
  axis_text_size = 3, 
  is_color = TRUE) +
  scale_color_manual(values = c("#999999",'#ff7f00')) +
  theme(
    legend.position = "none"
  )

mnist_proj_tsne_model1_link <- plot_proj(
    proj_obj = proj_obj1,
    point_param = c(1.2, 0.5), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-1.1, 0.9), 
    title = "b", 
    cex = 2, 
    axis_text_size = 3, 
    is_color = TRUE) +
    scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
    theme(
        legend.position = "none"
    )

@@

proj_obj2[["cluster"]] <- tsne_minst_scaled$cluster

mnist_proj_tsne_model2 <- plot_proj(
    proj_obj = proj_obj1,
    point_param = c(0.5, 0.3), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-1, 1), 
    title = "c", 
    cex = 1.5, 
    axis_text_size = 3, 
    is_color = TRUE) +
    scale_color_manual(values = c("#999999",'#ff7f00')) +
    theme(
        legend.position = "none"
    )

mnist_proj_tsne_model2_link <- plot_proj(
    proj_obj = proj_obj1, 
    point_param = c(1.2, 0.5), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-1, 1), 
    title = "c", 
    cex = 2, 
    axis_text_size = 3, 
    is_color = TRUE) +
    scale_color_manual(values = c("#d9d9d9",'#ff7f00')) +
    theme(
        legend.position = "none"
    )
